\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,hyperref}
\title{Primitive Polynomials as a Negligible Linear Slice of the Binary de Bruijn Space}
\author{Research addendum}
\date{\today}
\begin{document}
\maketitle

\begin{abstract}
Primitive polynomials over \(\mathrm{GF}(2)\) generate maximal-length LFSR sequences of period \(2^n-1\). After the usual zero-state insertion/modification, these give a structured family of binary de Bruijn cycles. We compare this linear-derived family with the total number of binary de Bruijn cycles and show that its fraction is bounded by a super-exponentially small quantity. In the sheaf--homology framework, primitive-polynomial constructions are linear global sections inside a vastly larger nonlinear global-section space.
\end{abstract}

\section{Counts}

The number of primitive polynomials of degree \(n\) over \(\mathrm{GF}(2)\) is
\[
P(n)=\frac{\varphi(2^n-1)}{n}.
\]
The number of binary de Bruijn cycles of order \(n\) is
\[
D(n)=2^{2^{n-1}-n}.
\]
Thus the fraction of primitive-polynomial-derived cycles is bounded by
\[
F(n)\le \frac{P(n)}{D(n)}
=
\frac{\varphi(2^n-1)/n}{2^{2^{n-1}-n}}.
\]
Since \(\varphi(2^n-1)\le 2^n-1<2^n\),
\[
F(n)\le \frac{2^{2n-2^{n-1}}}{n}.
\]
Therefore \(F(n)\to0\) super-exponentially.

\section{Interpretation}

A primitive polynomial gives a linear recurrence and therefore a linear global section of the word sheaf. The full de Bruijn space contains all global cyclic sections satisfying full coverage. Hence
\[
\text{primitive-polynomial-derived cycles}
\subset
\text{linear global sections}
\subset
\text{all de Bruijn cycles}.
\]
The asymptotic bound shows that almost all de Bruijn cycles are nonlinear.

\section{Homology}

A de Bruijn cycle is a full-support integral 1-cycle
\[
\partial c=0.
\]
Primitive-polynomial-derived cycles are those satisfying an extra linear recurrence. Thus primitive polynomials describe a small algebraically constrained region of the full integral cycle space.

\section{Conclusion}

Primitive polynomials provide efficient and elegant algebraic generators, but they do not represent the full de Bruijn space. The sheaf--homology framework explains this by distinguishing linear global sections from the vastly larger nonlinear global-section regime.
\end{document}
